---
title: "ESM 206 Lab 5"
author: "Crystal Bae"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives:

- Getting counts for different groups
- Use {lubridate} to parse dates
- Find confidence intervals & do t-tests with t.test()
- Heatmap with geom_tile()

```{r, include = FALSE}
library(tidyverse)
library(here)
library(janitor)
```

```{r}
# Import lobster data (lobster_abundance.csv) from SB long-term ecological survey
lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()

# Data is observed from 5 different sites
unique(lobster_abundance$site)
```

Use the tidyr::uncount() function to expand a frequency variable.

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count) # now each lobster has its own row
```

(Note: By default, tidyr::uncount() removed the "na" values.)

Now each measure is in its own cell (each lobster has its own row). This is tidy format.

Let's look at the data a bunch of different ways, using exploratory plots.

```{r}
# Jitterplot
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2)

# Few too many points to be useful...
```

```{r}
# Violin plot
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site))

# Same thing plotted as a violin plot, since we have a lot of observations.
```


