---
title: "ESM 206 Lab 5"
author: "Crystal Bae"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives:

- Getting counts for different groups
- Use {lubridate} to parse dates
- Find confidence intervals & do t-tests with t.test()
- Heatmap with geom_tile()

```{r, include = FALSE}
library(tidyverse)
library(here)
library(janitor)
```

## Import data and tidy up

```{r}
# Import lobster data (lobster_abundance.csv) from SB long-term ecological survey
lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()

# Data is observed from 5 different sites
unique(lobster_abundance$site)
```

Use the tidyr::uncount() function to expand a frequency variable.

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count) # now each lobster has its own row
```

(Note: By default, tidyr::uncount() removed the "na" values.)

Now each measure is in its own cell (each lobster has its own row). This is tidy format.

## Exploratory plots and check distributions

Let's look at the data a bunch of different ways, using exploratory plots.

```{r}
# Jitterplot
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2)

# Few too many points to be useful...
```

```{r}
# Violin plot
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site))

# Same thing plotted as a violin plot, since we have a lot of observations.
```

```{r}
# Histogram

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free") # to give each site its own plot

# by default each is given the same scale with facet_wrap(), but can add scales = "free" to give each own scale
# (of course, don't do that if you're COMPARING ACROSS groups)
# here we're just interested in the distribution WITHIN each group, to assess normality
```

```{r}
# Quantile-Quantile Plot (QQplot)
# Plots your sample values against the theoretical normal values

ggplot(data = lobster_tidy, aes(sample = size_mm)) + # need to specify which is the sample
  geom_qq() +
  facet_wrap(~site, scales = "free")
```

Tests for normality have their issues (tests w/ small samples often miss deviations from normality, while tests w/ large samples often falsely confirm deviations from normality), but you can look at these plots generally.

For the most part, these data look normally deviated for each group.

## Use 'lubridate' to parse dates and times

Since lubridate-type functions show up in various packages, use *lubridate::* to be specific.

Add a new column with *mutate()* that contains my date as an actual date:

```{r}
class(lobster_tidy$date) # currently date is just stored by R as a character column

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date) # our dates are in M/D/Y
  )

class(lobster_date$date_new)
```

**Note**: For storing dates in the first place, use ISO format; or better, store year/month/day all in separate columns.

Next, parse the date column to get the different pieces (month, year) separated.

```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_yr = lubridate::year(date_new),
    obs_mo = lubridate::month(date_new, label = TRUE), # label equals TRUE if you want it written out rather than the number
    obs_da = lubridate::day(date_new)
  )

levels(lobster_parse_date$obs_mo) # ordering of months is preserved! it's stored as an ordered factor
```

Count lobsters by different groupings

